---
title: "Demonstration of R Markdown"
author: "JAS"
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## R Markdown Demonstration using Breast Imaging Data
Data Citation: This breast cancer database was obtained from the University of Wisconsin Hospitals, Madison from Dr. William H. Wolberg.
See Machine learning techniques to diagnose breast cancer from image-processed nuclear features of fine needle aspirates. Cancer Letters 1994; 77:163-171.

***

### Step 1: Load Packages
We will be using two packages in addition to base R.
Amelia is for missing data and will provide us with a plot of missingness.
Caret is primarly from classification and regression trees, but has many features that are useful for data preprocessing. We will be using a function to partition data into training and testing.

```{r packages}
# Install and load required packages
# install.packages(c("Amelia", "caret", "tidyverse"))
library(Amelia)
library(caret)
library(tidyverse)

```

# Step 2: Load Data into R Environment and Perform Data Cleaning
# Illustrates loading a flat text file, assigning column names, checking variable types, recoding missing indicators, changing variable types, exploring variable distributions, creating a quick plot to indicate missingness, removing duplicates and missings, and creating a quick and dirty plot to compare features across outcome levels

```{r cleaning}

#Path where datasets are stored are stored as working directory
#setwd("C:/Users/js5406/OneDrive - cumc.columbia.edu/ML_Epi/2024/Data")

bc.data <- read.csv("./breast-cancer-wisconsin.data.txt", header = FALSE)

var.names <- c("id", "clump_thickness", "uniformity_csize", "uniformity_cshape", "marg_adhesion", "single_ecell_size", "bare_nuclei", "b_chromatin", "normal_nucleoli", "mitoses", "outcome")

colnames(bc.data) <- var.names

bc.data %>%
  str()

bc.data<-bc.data %>%
  mutate(across(everything(), ~ifelse(. == "?", NA, .))) %>%
  mutate(bare_nuclei = as.numeric(bare_nuclei),
         outcome = as.factor(outcome),
         outcome = fct_recode(outcome, "Benign" = "2", "Malignant" = "4")) 

bc.data %>% 
  summary() 

bc.data %>%
  missmap(main = "Missing values vs observed")

bc.data <- bc.data %>%
  na.omit()

bc.data <- bc.data %>%
  distinct(id, .keep_all = TRUE)

# Quick plot comparing clump thickness across Outcome Groups
bc.data %>%
  ggplot(aes(y = clump_thickness, x = outcome)) +
  geom_boxplot() +
  theme_classic()

bc.data$id <- NULL
```

# Step 3: Construct logistic regression models to predict Malignancy
# Model 1: Include all features
# Model 2: Include only clump thickness

```{r models}
model.1 <- glm(outcome ~ ., family = binomial(link = 'logit'), data = bc.data)
model.1 %>%
  summary() %>%
  print()
model.1 %>%
  confint() %>%
  print()

model.2 <- glm(outcome ~ clump_thickness, family = binomial(link = 'logit'), data = bc.data)
model.2 %>%
  summary() %>%
  print()
ci <- confint(model.2)
ci %>%
  print()

# Store results for later use--simple base stats way
OR <- exp(model.2$coefficients[2])
LL <- exp(ci[2, 1])
UL <- exp(ci[2, 2])

#Print model results
cat("OR:", OR, "(", LL, ",", UL, ")")

```




```

